---
title: "Campus Car Study"
output: html_document
date: "2026-09-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Is there a difference in status of cars driven at Adelphi between the faculty, resident and commuter students? 
## Hypothesis: We will see that the commuter students will have a higher proportion of status level cars, luxury cars, than faculty or resident students at Adelphi. 

Data Collection Methods: 
Explanatory Variable: type of parking sticker (commuter, resident, faculty)
Response Variable: brand of car
Covariate: open lots versus garage, time of day, and day of the week 
When: data will be collected on a weekday and mid-morning to capture all three populations
Where: Nexus lot (1) and garage, then science building lot (4) and garage
Sample size: in each parking lot, we will haphazardly identify 9 of each sticker type (27 cars per 4 lots; n = 108). 


```{r}

library(readxl)
library(dplyr)
##read the excel files so the two sheet tabs can be read
##separate import by sheet tab
adelphidata <- read_excel('/Users/meganyoumans/Desktop/Adelphi Coursework/Statistical Modeling/Assignments/HW 4/Adelphi Car Sticker Dataset.xlsx', sheet = "Data")
carprices <- read_excel('/Users/meganyoumans/Desktop/Adelphi Coursework/Statistical Modeling/Assignments/HW 4/Adelphi Car Sticker Dataset.xlsx', sheet = "Mean")
alldata <- adelphidata %>% left_join(carprices, by = "Brand")
##join both data sets so that there is a read for all of the data all at once to be able to compare the mean prices of brand of cars 

names(adelphidata)
names(carprices)

summary(adelphidata)
summary(carprices)

```

## Plots

```{r, echo=FALSE}

carcount <- table(alldata$`Sticker Type`, alldata$Brand)
barplot(carcount, col = c("yellow", "brown", "gray"),
        legend.text = TRUE,
        args.legend = list(x="top"),
        las = 2, ##rotates axis labels
        main = "Car Brand Counts",
        ylab = "Count")

sorted <- carprices[order(carprices$`Mean Cost (Numerical)`),]
barplot(sorted$`Mean Cost (Numerical)`,
        names.arg = sorted$Brand,
        col = "yellow",
        main = "Mean Cost of Car Brands")


carcount <- table(alldata$Observers, alldata$Brand)

barplot(carcount,
        col = c("yellow", "brown", "gray"),
        legend.text = rownames(carcount),
        args.legend = list(title = "Observers", x = "top"),
        las = 2,
        cex.names = 0.7,
        ylab = "Count",
        main = "Brand by Observer")

mosaicplot(carcount,
           color = rainbow(ncol(carcount)),
           las = 2,
           main = "Observer vs. Brand",
           xlab = "Observer",
           ylab = "Brand")

chisq.test(carcount)


##Are Observers biased to type of car? 

##make a car count table
carcount <- table(alldata$Observers, alldata$Brand)
## row percentages
carpercent <- prop.table(carcount, margin=1) * 100
## prop table turns counts into proportional data decimals between 0 and 1 
##margin = 1 how to calculate proportions in each row to sum to 1 or 100% 
## *100 takes the decimals and turns them into percentags 

rowSums(carpercent)
##checks percent

##use t. to flip row and columns for the desired bar plot
t(carpercent)

##set space for the margins for the larger legend
par(mar = c(8,4,4,10))

barplot(t(carpercent),
        beside = TRUE, ##grouped bars by observer
        col = rainbow(ncol(carpercent)), ##one color per brand
        las = 2, ##rotates the text
        cex.names = 0.8, ##cex makes text size smaller 
        ylab = "% Oberservers Brands",
        main = "Brands by Observers (%)")
legend (x="topright", inset = c(-0.17, 0), ## negative pushes it outside of box
        legend = colnames(carpercent),
        fill = rainbow(ncol(carpercent)),
        title = "Brand",
        cex = 0.6,
        xpd = TRUE)

## car type to lot seen 

carcount2 <- table(alldata$Lot, alldata$Brand)
par(mar = c(8,4,4,10))

barplot(t(carcount2),
        beside = TRUE, ##grouped bars by observer
        col = rainbow(ncol(carcount2)), ##one color per brand
        las = 2, ##rotates the text
        cex.names = 0.8, ##cex makes text size smaller 
        ylab = "% Lot Brand",
        main = "Brands by Lot (%)")
legend (x="topright", inset = c(-0.17, 0), ## negative pushes it outside of box
        legend = colnames(carpercent),
        fill = rainbow(ncol(carpercent)),
        title = "Brand",
        cex = 0.6,
        xpd = TRUE)

##One group, Alyssa and Danielle, saw way more toyota than the rest of the groups 
##More Toyota were seen in the SCB lot than anywhere else, then an Nissan and Honda in the middle ground, with a low distribution of most other car brands everywhere 
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
