---
title: "Argiope Small and Quine Megan Youmans"
output:
  word_document: default
  html_document: default
date: "2026-09-22"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


spiders <- read.csv('/Users/meganyoumans/Desktop/Adelphi Coursework/Statistical Modeling/Datasets 2/Argiope_fertilization_small.csv')

##SMALL ARGIOPE DATA SET 

quine <- read.csv('/Users/meganyoumans/Desktop/Adelphi Coursework/Statistical Modeling/Datasets 2/quine.csv')

## QUINE DATA SET 


```

## Argiope 

For the Agriope small fertilization dataset, do multivariate linear models to
test if copulation insertion duration correlates to the number of fertilized eggs.
To do this assessment of male fertility, we must also account for male size, female size, eggs
laid, and egg sacs produced


```{r agriope}
summary(spiders)

spiderlm <- lm(data=spiders, spiders$NumFertilized ~ spiders$InsertionDuration + spiders$MaleSize + spiders$FemaleSize + spiders$NumSacs + spiders$NumEggs)
summary(spiderlm)

```

## Significant p-values notes for number of eggs correlating to number of eggs fertilized, as well as inderstion duration to number of eggs fertilized. Therefore we can assume that if a spider female has more eggs, and the insertion duration for the male is lnog that there is a likelihood of a higher number of eggs being fertilized. 



```

```


## Quine

do multivariate categorical ANOVAs and Tukey’s
Honest Significant Differences tests to determine if the number of days absent is predicted by
Ethnicity, Sex, Age, or Learner Status


```{r quine, echo=FALSE}
summary(quine)

library(ggplot2)

ggplot(quine) + aes(x = Sex, y = Days, fill = Eth) + geom_boxplot()

ggplot(quine) + aes(x = Sex, y = Days, fill = Age) + geom_boxplot()

ggplot(quine) + aes(x = Sex, y = Days, fill = Lrn) + geom_boxplot()

mod <- aov(quine$Days ~ quine$Eth + quine$Sex + quine$Age + quine$Lrn)
summary(mod)

ggplot(quine) + aes(x = Age, y = Days, fill = Eth) + geom_boxplot()

##p-values for Ethnicity and age showing strong correlations respectively for how many days absent where Eth "A" and older age show a higher correlation of missing days in school, although Eth "N" has outliers to the norm as well. 

mod <- aov(quine$Days ~ quine$Eth + quine$Sex + quine$Age + quine$Lrn)
summary(mod)
##Tukey test, using aov(), comparison of means for confidence levels
TukeyHSD(mod)

##p adjusted for Eth is 0.0004658
##P sex adjusted is not significant, gender was not a major influence on missing days
##P value for F2-F1 value is 0.045, slightly older in age is significant for missing days versus other age groups
##P value for learner status is not significant for missing days 

##Ethnicity differences played a role in students missing school days, as well as age groups F2 and F1 age groups were more significant for missing days than other students. 


```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
