---
title: "Wheat Megan Youmans"
output:
  html_document: default
  word_document: default
date: "2026-09-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## R Markdown

##Wheat Data Set
## yield, winter, ears, pH, P, K, Mg 

##Hypothesis
##Variables are all numerical 
##yield of wheat is the response to the environmental factors, predictors could be winter(temperature), ears and pH of soil
##co variates could be the quality of the soil with P, K, and Mg levels 

##Seeing correlations between P and pH levels, and also Yield to Mean number of Overwintering plants. 
## Plotting how P and pH values compared against yield and overwintering mean amounts 


```{r wheat}
wheat <- read.csv('/Users/meganyoumans/Desktop/Adelphi Coursework/Statistical Modeling/Datasets 2/wheat.csv')
wheat
summary(wheat)

```

## Plots

```{r, echo=FALSE}
plot(wheat)
wheat_corr <- cor(wheat[,2:7])
wheat_corr
corrplot::corrplot(wheat_corr, method = "color")
## pH and P correlation noted 
car::scatterplotMatrix(~wheat$pH+wheat$P+wheat$yield, id=list(n=3), data=wheat)
## pH increases, P increases, but as P increases, wheat yield potentially decreases, but midlevel pH seems to be best for wheat? 

##scatterplots 
##had an issue with the scatterplotMatrix and scatterplots with car data set for some reason, adding car:: was the workaround 
car::scatterplot(wheat$yield,wheat$ears)
car::scatterplot(wheat$P,wheat$pH)
car::scatterplot(wheat$yield,wheat$ears)
car::scatterplot(wheat$winter,wheat$yield)

##amount of overwintering plants survive based on P and pH  

wheatoverwinteringplants <- cut(wheat$winter, breaks = 3, labels = c("Few","Medium","Most")) 
##separate out amount of overwintering plants counted 

par(mar = c(5,4,4,8))
## add space outside box for legend

interaction.plot(
  x.factor=wheat$pH,
  trace.factor = wheatoverwinteringplants,
  response = wheat$P,
  fun = mean,
  type = "b",
  col=1:3, pch=1:3,
  xlab="pH",
  ylab="Soil P",
  main = "Soil Factors",
  trace.label = "Mean # of Overwintering Plants",
  legend = FALSE
)
legend("bottomright", inset = c(-0.30,0), xpd=TRUE, title="Mean # of Overwintering Plants", leg.txt <- c("Few", "Medium", "Most"), col=1:3,pch=1:3)

## best average means of overwintering plants counted between pH 5.6-6.4, P levels 60-90 


## Wheat yield affected by P and pH 

wheatyieldamount <- cut(wheat$yield, breaks = 3, labels = c("Few","Medium","Most")) 
##separate out amount of yield  

par(mar = c(5,4,4,8))
## add space outside box for legend

interaction.plot(
  x.factor=wheat$pH,
  trace.factor = wheatyieldamount,
  response = wheat$P,
  fun = mean,
  type = "b",
  col=1:3, pch=1:3,
  xlab="pH",
  ylab="Soil P",
  main = "Soil Factors",
  trace.label = "Wheat Yield Amount",
  legend = FALSE
)
legend("bottomright", inset = c(-0.30,0), xpd=TRUE, title="Wheat Yield Amount", leg.txt <- c("Few", "Medium", "Most"), col=1:3,pch=1:3)

##wheat yield is highest between pH 6 and 6.4 and P levels at around 60
##

##

plevels <- cut(wheat$P, breaks = 2, labels = c("Low", "High")) 
pHlevels <- cut(wheat$pH, breaks = 2, labels = c("Low", "High")) 

wheattrace <- interaction(plevels, pHlevels, sep = "& pH")

par(mar = c(5,4,4,12))
## add space outside box for legend

interaction.plot(
  x.factor=wheat$winter,
  trace.factor = wheattrace,
  response = wheat$yield,
  fun = mean,
  type = "b",
  col=1:6, pch=1:6,
  xlab="Winter",
  ylab="Yield",
  main = "Result",
  trace.label = "P & pH Levels",
  legend = FALSE
)

legend("bottomright", inset = c(-0.45,0), xpd=TRUE, title="P & pH", legend = levels(wheattrace), col=1:6,pch=1:6)

## best Yield to lowest mean avg of overwintering plants tends to be with high P and low pH 


##boxplot 
boxplot(wheat$yield~wheat$winter)
boxplot(wheat$yield~wheat$P)
##wheat more tolerant at low to medium P levels  
boxplot(wheat$yield~wheat$pH)
##wheat seems tolerant of pH 5.4-6.4


```


